%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE acknowledgements.tex}

\begin{ntacknowledgements}

I would like to express my gratitude to Professor Mauro Castelli for all the support, understanding and advice given throughout this process. Thank you for the words of encouragement and the outstanding availability and reliability as an advisor.

I am grateful to the Alfred Wegener Institute and Dr. Ingmar Nitze for providing me with the labelled data that made this project possible. I would also like to thank Dr. Nitze for the continued guidance and support in the unknown world of remote sensing imagery and permafrost.

I would also like to thank my leaders at Accenture, Fernando Lucini and Gilbert Cassar for all the support and flexibility shown to me this past year, while I wrote this dissertation still working full-time. A special thanks to Gilbert for all the advice and technical knowledge shared with me during the writing process. I would also like to thank Adam Wickened for helping me with the input data collection and also for going through ideas with me on model development. Finally, to Jeffrey Nogosek for organising Cloudwars 3 with the Alfred Wegener Institute and the Stockholm Environment Institute, without this project, I would not have had the awareness, data or understanding needed to pursue this work.

A special thanks to my grandmother, mum, dad, auntie and uncle who have opened every door and window they could to give me all the oportunities that have led me where I am today. Thank you for respecting and supporting all my decisions and all the love you give me uncondicionally. To my brother, for being my best friend and always there for me in times of need, always available to help without any questions. To Alex, for being my partner in crime, for keeping me sane and grounded through this process and supporting me in every way she could. To Moyin, Helen and Annie for keeping me on track, helping me break this down into small goals with Trello and  for always making the time every week to check on me and cheer me on. To Charlotte for our writing club and all the advice and tips. Thank you to Carlos for helping me at very short notice with my very rusty Portuguese. Thank you to Francesca, Charlotte, Chuck, Gabriel, Kristia, Tze Ni, SiSi, Alex, Catarina, Carolina, Mafalda, Miguel, Maren, Manel, Mariana, Inês, Jona, Hannah, Joey and Tugse for all the kind words of encouragement and all the advice you gave me throughout this process.

I can not thank enough everyone that has been mentioned and so many other friends that have shown me nothing but kindness and support. Thank you for believing in me and thank you for helping me grow everyday.

This was an incredible journey to be on, it was tough, it was outside of my comfort zone, but in the end it was worth it. Thinking of that blank page and seeing how far I was able to come with everyone's support makes me be hopeful and positive about all the things we can achieve together. Finally, I am thankful for hope.
\end{ntacknowledgements}